Search-Based Test Data Generation for JavaScript Functions that Interact with the DOM \thanks{Identify applicable funding agency here. If none, delete this.}

### 2018-05-14 doc/issre2018/introduction.tex
Final edits of the paper

According to the \emph{StackOverflow} Developer Survey 2018 \cite{stackoverflow2018}, JavaScript has been recognized for the sixth time in row as the most commonly used programming language. It is also the most popular language on \emph{GitHub}~\cite{guthub2017} based on the number of opened pull requests (2.3 million). Such an ever increasing JS success can be explained by its native browsers support, and the rise in popularity of the JS run-time environment \emph{Node.js}. On the other hand, due to of implicit type conversion, lack of static types, and extensive support of dynamic language constructs~\cite{richards2010analysis}, JS applications are often error-prone~\cite{frolin:TSE16}, hard to analyze~\cite{andreasen2017survey, sun2017analysis} and test~\cite{mesbah2015advances}. Thus, it is not surprising that once in a while critical bugs manage to break through~\cite{bugstories2017} into the web applications by Facebook, Twitter etc.

Systematic unit testing is basic and, at the same time, the most effective technique for improving software quality. In practice, there are numerous of JS frameworks --- Mocha, Jasmine, Jest, Enzyme, Tape, Ava, Karma, QUnit etc. --- intended to assist developers with the automation of various unit testing activities such as design, execution and monitoring. Despite the apparent abundance of the JS testing tools, the developers' satisfaction rate is still relative low~\cite{stateJS2017} (3.2 out of 5), suggesting the need for improvements. One of such areas is \emph{test data generation}. It pursues the goal of maximizing the test coverage by exercising the function under test (FUT) with different input parameters. Moreover, if a FUT interacts with the DOM of a web application, a \emph{test fixture} is expected as an additional input.

Generally, the problem of test data generation has been under an active study for the past three decades. In particular for JS, researchers have proposed several solutions based on: concolic execution (\emph{Jalangi}~\cite{sen2013jalangi}, \emph{Confix}~\cite{amin:ase15}), random generation (\emph{JSContest}~\cite{heidegger2010contract}), and the analysis of the whole web application (\emph{Artemis}~\cite{artemis2011}, Crawljax~\cite{mesbah2012crawling}). Out of those tools, only \emph{Confix} tries to construct a sufficient DOM fixture by translating native browser API calls into logical constraints. However, in the empirical evaluation~\cite{amin:ase15}, Confix only reached about 50\% in \emph{branch coverage} on the selected subjects. At the same time, search-based techniques have shown prominent results in achieving high branch coverage for imperative~\cite{wegener2001evolutionary}, object-oriented~\cite{fraser2011evosuite} and dynamic languages~\cite{irawan2016test, wibowo2015unit}, as well as for web~\cite{alshahwan2011automated} and mobile~\cite{mao2016sapienz} applications. But to the best of our knowledge, the problem of test data generation has never been tackled for JS with the help of the search-based approach. In this paper we are breaching this gap by introducing a \textbf{J}avascript \textbf{E}volution-based testing framework with \textbf{D}OM as an \textbf{I}nput, called \emph{JEDI}. We can summarize our contributions as follows:
\begin{enumerate}[leftmargin=5mm]
\item The \emph{JEDI} framework is a novel JS unit testing tool using search-based techniques for generating test data. Notably, our testing framework is able to generate an arbitrary DOM input (both tags and attributes), which is syntactically valid and compliant with the recent HTML5 specification.
\item Our test generation algorithm, called ``genetic with restart'', is trying to escape search plateau by restarting GA with a new target. Because the full-fledged JS data flow analysis~\cite{jang2009points} is hard to reach, the new target is chosen purely based on the preceding control flow behaviour in contrast with more popular techniques~\cite{ferguson1996chaining}.
\item Based on the case studies found in two related works~\emph{Confix}~\cite{amin:ase15} and \emph{TAJS}~\cite{dom2011}, we conducted an empirical validation of our framework and performed a significance study of the results. The validation has shown the effectiveness of the framework in covering target branches with a reasonable efficiency. Moreover, we showed that on the selected subjects \emph{JEDI} outperforms \emph{Confix} ....
\end{enumerate}

The paper is structured as follows: In Section~\ref{sec.example}, we present a motivating example illustrating the challenges of generating test data for JS. The architecture of the \emph{JEDI} testing framework is explained in Section~\ref{sec.framework}. Section~\ref{sec.evaluation} presents empirical validation and indicates the validity threats. Related and future work are discussed in Section~\ref{sec:related.work}. Section~\ref{sec:concl} concludes the paper.

>>>>

According to the \emph{StackOverflow} Developer Survey 2018 \cite{stackoverflow2018}, JavaScript has been recognized for the sixth time in row as the most commonly used programming language. It is also the most popular language on \emph{GitHub}~\cite{guthub2017} based on the number of opened pull requests (2.3 million). The success of JS can be explained by its native browsers support, and the rise in popularity of the JS run-time environment \emph{Node.js}. On the other hand, due to of implicit type conversion, lack of static types, and extensive support of dynamic language constructs~\cite{richards2010analysis}, JS applications are often error-prone~\cite{frolin:TSE16}, hard to analyze~\cite{andreasen2017survey, sun2017analysis} and test~\cite{mesbah2015advances}. Thus, it is not surprising that once in a while critical bugs manage to break through~\cite{bugstories2017} even into the web applications of major companies such as Facebook, Twitter etc.

Systematic unit testing is a basic and, at the same time, the most effective technique for improving software quality~\cite{martin2009clean}. There exist numerous JS frameworks on the market --- Mocha, Jasmine, Jest, Enzyme, Tape, Ava, Karma, QUnit etc. --- intended to assist web developers with the automation of various  activities such as the design, execution and monitoring of unit tests. Despite the apparent abundance of the JS testing tools, the developers' satisfaction rate is still relative low~\cite{stateJS2017} (3.2 out of 5), suggesting the need for improvements. One of the missing steps on the way to complete JS test automation is \emph{the generation of test data}. It pursues the goal of maximizing test coverage by exercising the function under test (FUT) with various input parameters. Moreover, if a FUT interacts with the DOM of a web application, a \emph{test fixture} is expected as an additional input.

Generally, the problem of test data generation has been actively studied for the past three decades. In particular for JS, researchers have proposed several solutions based on: concolic execution (\emph{Jalangi}~\cite{sen2013jalangi}, \emph{Confix}~\cite{amin:ase15}), random generation (\emph{JSContest}~\cite{heidegger2010contract}), and the analysis of the whole web application (\emph{Artemis}~\cite{artemis2011}, Crawljax~\cite{mesbah2012crawling}). Out of those tools, only \emph{Confix} tries to construct sufficient DOM fixtures by encoding browser API calls as logical constraints. However, the empirical validation of Confix~\cite{amin:ase15} has shown that it only reached about 50\% \emph{branch coverage} on the selected subjects. At the same time, search-based techniques have shown prominent results in successfully achieving higher branch coverage for imperative~\cite{wegener2001evolutionary}, object-oriented~\cite{fraser2011evosuite} and dynamic languages~\cite{irawan2016test, wibowo2015unit}, as well as for web~\cite{alshahwan2011automated} and mobile~\cite{mao2016sapienz} applications. But to the best of our knowledge, the problem of test data generation has never been tackled for JS with the help of the search-based approach. In this paper we are breaching this gap by introducing a \textbf{J}avascript \textbf{E}volution-based testing framework with \textbf{D}OM as an \textbf{I}nput, called \emph{JEDI}. We can summarize our contributions as follows:
\begin{enumerate}[leftmargin=5mm]
\item The \emph{JEDI} framework is a novel JS unit testing tool using search-based techniques for generating test data. Notably, our testing framework is able to generate an arbitrary DOM input (both tags and attributes), which is syntactically valid and compliant with the recent HTML5 specification.
\item Our test generation algorithm, called ``genetic with restart'', attempts to escape a plateau by restarting the search with a new target. In contrast with the \emph{chaining approach}~\cite{ferguson1996chaining}, we choose the new target purely based on the preceding control flow behaviour because a full-fledged data flow analysis~\cite{jang2009points} of JS is hardly reachable.
\item Based on the case studies found in two related works~\emph{Confix}~\cite{amin:ase15} and \emph{TAJS}~\cite{dom2011}, we conducted an empirical validation of our framework and performed a significance study of the results. The validation has shown the effectiveness of the framework in covering target branches with a reasonable efficiency. Moreover, we show that out of 10 functions where \emph{JEDI} achieved complete branch coverage, \emph{Confix} only covered 32\% on average.
\end{enumerate}

The paper is structured as follows: In Section~\ref{sec.example}, we present a motivating example illustrating the challenges of generating test data for JS. The architecture of the \emph{JEDI} testing framework is explained in Section~\ref{sec.framework}. Section~\ref{sec.evaluation} presents empirical validation and the validity threats. Related and future work are discussed in Section~\ref{sec:related.work}. Section~\ref{sec:concl} concludes the paper.

### 2018-05-14 doc/issre2018/related-work.tex
Final edits of the paper

The last two decades have been highly productive for the research in JS testing and analysis~\cite{andreasen2017survey, sun2017analysis, mesbah2015advances}. We discuss the state-of-the-art in JS testing clustering the works by the underlying technique: feedback-directed random generation~\cite{heidegger2010contract,heidegger2012jscontest,kristensen2017type}, symbolic~\cite{tanida2014automatic,saxena2010symbolic} and concolic~\cite{sen2013jalangi,amin:ase15} execution, static~\cite{tajs2009,dom2011} and dynamic~\cite{andreasen2017survey} analysis, formal verification~\cite{gardner2008local,lerner2012modeling,gardner2012towards}, search-based testing~\cite{alshraideh2008complete}, and the whole web application testing~\cite{alshahwan2011automated,mesbah2012invariant,milani2014leveraging}.

Guided by the developer's provided type annotation, \emph{JSContest}~\cite{heidegger2010contract} generates random input for a JS function with the goal to verify the conformance of \emph{type} or \emph{access permission}~\cite{heidegger2012jscontest} contracts. It constructs test inputs out statically collected constants. In comparison, our genetic-based approach can handle interactions with the DOM, and also leverages dynamic information. $TS_{TEST}$~\cite{kristensen2017type} is the feedback-directed random testing tool for checking the conformance of TypeScript declaration files with the respective JS libraries. It supports generics but still lacks DOM generation.

@@

\emph{Jalangi}~\cite{sen2013jalangi} is a JS dynamic analysis framework which can be instantiated for \emph{concolic testing} or {type inconsistency} analysis~\cite{pradel2015typedevil} but it does not DOM. \emph{Confix}~\cite{amin:ase15} translates DOM-related statements into XPath expressions, and feeds them to a constrain solver in construct DOM fixtures. However, it could reinforce \Jalangi with the capability of DOM modeling. As our evaluation has shown, \Confix can not find solutions for programs with complex data dependency. SymJS~\cite{tanida2014automatic} employs symbolic execution for JS unit-test generation but again missing symbolic stubs for DOM API.

@@

Jensen et al.~\cite{tajs2009} proposed a static analysis framework \emph{TAJS} for JS type inference. Later they extend the analysis to support reasoning about HTML DOM and browser API~\cite{dom2011}. \emph{TAJS} is able to point to a dead code, type violation and references to \emph{undefined} values.

Lerner~\cite{lerner2012modeling} define formal model of event behavior in the DOM to establish testing oracles. Gardener et al.~\cite{gardner2012towards} tries to esteblish formal versification framework for JS including part of the DOM specification~\cite{gardner2008local}.

Alshraideh~\cite{alshraideh2008complete} is the only attempt to employ GA for JS test data generation. But the work deals with the functions of primitive type, without DOM, and has no not been empirically validated. Recently, SBST was applied to other dynamic languages such as Ruby~\cite{mairhofer2011search}, Lua~\cite{wibowo2015unit} and Go~\cite{irawan2016test}, but the treatment of DOM still makes our approach unique. DOM is an instance of \emph{tree-like} data structure. \emph{XMLMATE}~\cite{havrikov2014xmlmate} is an XML test generation tool build on top \emph{EvoSuite} with the aim of improving test coverage of the XML-aware Java programs. To generate a random document, it either relies on the  existing XML schema, or generates the document from scratch. In some sense, DOM specification can be thought of a complex XML schema, with extensive validation requirements. One distinction of our approach is the use of program analysis for data generation. Both tools have similar philosophy for the mutation and crossover. Jan et al.~\cite{jan2016automated} suggest to use XML fuzzing for identifying vulnerabilities in web services.

In contrast with the early presented techniques, the whole web application testing requires the application to be fully functional. This enables test generation based on crawling~\cite{mesbah2012invariant} and can be combined with the developer's written tests~\cite{milani2014leveraging}. \emph{Artemis}~\cite{artemis2011} is a feedback-directed random testing tool for JavaScript web application. It discovers new tests by generating sequences of executable events and monitoring they effect on the state of the application. Biagiola et al.~\cite{biagiola2017search} proposed to use \emph{page objects} to generate navigation path and input data with by exploiting search-based approach. \emph{JSEFT}~\cite{mirshokraie2015jseft} leverages dynamic execution to generate JS unit test including fixtures captured at run-time. The generated test can be compeented with DOM-dependent assertions produced by \emph{Atrina}~\cite{icst16}.

\emph{Flycatcher}~\cite{de2012automatic} generates tests by instrumenting a program with the help the JS \emph{Proxies} in order to collect type related information.

\emph{$JS_{DEP}$}~\cite{sung2016static} proposes a constraint-based static analysis for learning dependencies between DOM-events, which can help to prune the sequence of tests consisting of the web-application events.

@@

One likely direction for the future extension of \emph{JEDI} is \emph{the whole test suite generation}~\cite{fraser2013whole}. Such problems usually incur multi-objective optimization~\cite{lakhotia2007multi,panichella2017lips}, e.g. trade-off between coverage and test-suite length.

Another direction of future work is re-targeting out tool towards mutation-driven generation of unit tests and oracles, similar to the line of work~\cite{fraser2012mutation}. That should also help to evaluate fault revealing capability of \emph{JEDI} as the number of killed mutants.

>>>>

The last two decades have been highly productive for the research in JS testing and analysis~\cite{andreasen2017survey, sun2017analysis, mesbah2015advances}. Discussing the state-of-the-art in JS testing we cluster the works, by the underlying technique, into the following categories: feedback-directed random generation~\cite{heidegger2010contract,heidegger2012jscontest,kristensen2017type}, symbolic~\cite{tanida2014automatic,saxena2010symbolic} and concolic~\cite{sen2013jalangi,amin:ase15} execution, static~\cite{tajs2009,dom2011} and dynamic~\cite{andreasen2017survey} analysis, formal verification~\cite{gardner2008local,lerner2012modeling,gardner2012towards}, search-based generation~\cite{alshraideh2008complete}, and whole web application testing~\cite{alshahwan2011automated,mesbah2012invariant,milani2014leveraging}.

Guided by developer's type annotations, \emph{JSContest} generates random inputs for JS functions with the goal to verify the conformance to the \emph{type}~\cite{heidegger2010contract} and \emph{access permission}~\cite{heidegger2012jscontest} contracts. In comparison with our genetic approach, it does not leverage dynamic information to construct test inputs, only static one, and ignores the DOM interactions. $TS_{TEST}$~\cite{kristensen2017type} employs the feedback-directed random generation for conformance checking of TypeScript declarations with the respective JS libraries. It supports generics, but cannot generate DOM.

@@

\emph{Jalangi}~\cite{sen2013jalangi} is a JS dynamic analysis framework which can be instantiated for \emph{concolic testing} and \emph{type inconsistency analysis}~\cite{pradel2015typedevil}, but it does not support DOM. \emph{Confix}~\cite{amin:ase15} translates DOM-related statements into XPath expressions, and feeds them to a constraint solver in order to construct a DOM fixture. It can reinforce \Jalangi with the capability of DOM modeling. However, as our evaluation has shown, \Confix cannot resolve hard to cover branches. \emph{SymJS}~\cite{tanida2014automatic} employs symbolic execution for JS unit-test generation but again has no symbolic stubs for the DOM API.

@@

Jensen et al.~\cite{tajs2009} proposed a static analysis framework \emph{TAJS} for JS type inference. Later the analysis was extended to support reasoning about the HTML DOM and browser API~\cite{dom2011}. \emph{TAJS} is able to point to a dead code, type violation and references to \emph{undefined} values. A review of dynamic analysis techniques ofr JS can be found in a recently published survey~\cite{andreasen2017survey} by Andreasen et al.

Lerner~\cite{lerner2012modeling} formally models the event behavior in the DOM to establish testing oracles. Gardener et al.~\cite{gardner2012towards} presents a formal verification framework for JS, which can deal with some parts the DOM specification~\cite{gardner2008local}.

Alshraideh~\cite{alshraideh2008complete} is the first work suggesting to use GA for JS test data generation. But the implementation is limited to primitive types such as numbers and strings, and does not report on empirical validation. Moreover, it does not support the generation of arrays and DOM. Recently, SBST has been applied to other dynamic languages such as Ruby~\cite{mairhofer2011search}, Lua~\cite{wibowo2015unit} and Go~\cite{irawan2016test}, but the treatment of DOM still makes our approach unique. \emph{XMLMATE}~\cite{havrikov2014xmlmate} is the only search-based tool we have encountered that can produce \emph{tree-like} structures such as XML. It is built on top of \emph{EvoSuite} with the aim of improving the test coverage of the XML-aware Java programs. To generate a random document, it either relies on an existing XML schema, or generates the document from scratch. Both tools have a similar approach for the mutation and crossover. One distinguishing feature of our approach is the use of program analysis to facilitate DOM generation. Jan et al.~\cite{jan2016automated} suggest to use XML fuzzing for identifying vulnerabilities in web services.

In contrast with the techniques discussed so far, whole web application testing requires the application to be fully functional. This enables test generation based on crawling~\cite{mesbah2012invariant} and can be combined with the developer's written tests~\cite{milani2014leveraging}. \emph{Artemis}~\cite{artemis2011} is a feedback-directed random testing tool for JavaScript web applications. It discovers new tests by generating the sequences of executable events and monitoring how they effect the application state. Biagiola et al.~\cite{biagiola2017search} propose to use existing \emph{page objects} to produce navigation paths and input data by exploiting the search-based approach. \emph{JSEFT}~\cite{mirshokraie2015jseft} leverages dynamic execution to generate JS unit test including fixtures captured at run-time. The generated test can be complemented with the DOM-dependent assertions produced by \emph{Atrina}~\cite{icst16}.

%% \emph{Flycatcher}~\cite{de2012automatic} generates tests by instrumenting a program with the help the JS \emph{Proxies} in order to collect type related information.

%% \emph{$JS_{DEP}$}~\cite{sung2016static} proposes a constraint-based static analysis for learning dependencies between DOM-events, which can help to prune the sequence of tests consisting of the web-application events.

@@

One likely direction for the future extension of \emph{JEDI} is \emph{whole test suite generation}~\cite{fraser2013whole}. Such problems usually incur multi-objective optimization~\cite{lakhotia2007multi,panichella2017lips}, e.g. a trade-off between coverage and test-suite length.

Another direction of future work is re-targeting our tool towards mutation-driven generation of unit tests and oracles, similar to~\cite{fraser2012mutation}. That should also help to evaluate fault revealing capability of \emph{JEDI} as the number of killed mutants.
